Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-21 paper.qmd
Changed the plot (a) for Figure 9

  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2

  scurve_model <- fit_highd_model(
    training_data = training_data_scurve,
    emb_df = umap_scurve_scaled,
    bin1 = xbins,
    r2 = r2,
    q = 0.07, 
    is_bin_centroid = TRUE,
    is_rm_lwd_hex = FALSE,
    col_start_highd = "x"
  )

@@

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve,
    df_bin = df_bin_scurve,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(bin1 = xbins,
           bin2 = bin2,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve))

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = bin1, 
                         y = MSE)) + 
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) + 
  # geom_vline(xintercept = 11, linetype="solid",
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 9, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 13, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  scale_x_continuous(breaks = 5:19) +
  labs(x = expression(b[1]), y = "MSE") +
  ggtitle("(b)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))

b1_m_scurve <- ggplot(error_scurve, 
                     aes(x = bin1, 
                         y = b_non_empty/b)) + 
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) + 
  # geom_vline(xintercept = 11, linetype="solid",
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 9, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 13, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  scale_x_continuous(breaks = 5:19) +
  labs(x = expression(b[1]), y = expression(frac(m, b))) +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))

b1_m_scurve2 <- ggplot(error_scurve, 
                     aes(x = bin1, 
                         y = b_non_empty)) + 
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) + 
  # geom_vline(xintercept = 11, linetype="solid",
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 9, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 13, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  scale_x_continuous(breaks = 5:19) +
  labs(x = expression(b[1]), y = expression(m)) +
  ggtitle("(c)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))
```

@@

b1_m_scurve + mse_scurve_b + b1_m_scurve2 + 
mse_scurve_lwd + mse_scurve_lwd2 +
distance_hist + plot_layout(ncol=2)
```

>>>>

  bin_obj <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)

  bin2 <- bin_obj$bin2
  a1 <- bin_obj$a1

  scurve_model <- fit_highd_model(
    training_data = training_data_scurve,
    emb_df = umap_scurve_scaled,
    bin1 = xbins,
    r2 = r2,
    q = 0.1, 
    is_bin_centroid = TRUE,
    is_rm_lwd_hex = FALSE,
    col_start_highd = "x"
  )

@@

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve,
    df_bin = df_bin_scurve,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(bin1 = xbins,
           bin2 = bin2,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve),
           a1 = a1)

@@

diaconis_bin_width <- calc_bins_y(
    bin1 = ceiling(NROW(training_data_scurve)^(1/3)),
    r2 = r2,
    q = 0.1)$a1

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = a1, 
                         y = MSE)) + 
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) + 
  # geom_vline(xintercept = 11, linetype="solid",
  #            color = "#bdbdbd", linewidth=1) +
  geom_vline(xintercept = diaconis_bin_width, 
             linetype=2, color = "#bdbdbd", 
             linewidth=1) +
  # geom_vline(xintercept = 13, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  #scale_x_continuous(breaks = 5:19) +
  labs(x = expression(a[1]), y = "MSE") +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))

# b1_m_scurve <- ggplot(error_scurve, 
#                      aes(x = bin1, 
#                          y = b_non_empty/b)) + 
#   geom_point(size = 1) +
#   geom_line(linewidth = 0.05) + 
#   # geom_vline(xintercept = 11, linetype="solid",
#   #            color = "#bdbdbd", linewidth=1) +
#   # geom_vline(xintercept = 9, linetype=2,
#   #            color = "#bdbdbd", linewidth=1) +
#   # geom_vline(xintercept = 13, linetype=2,
#   #            color = "#bdbdbd", linewidth=1) +
#   scale_x_continuous(breaks = 5:19) +
#   labs(x = expression(b[1]), y = expression(frac(m, b))) +
#   ggtitle("(a)") +
#   theme(aspect.ratio = 0.75,
#         axis.text.x = element_text(size = 12),
#         axis.text.y = element_text(size = 12),
#         axis.title.x = element_text(size = 12),
#         axis.title.y = element_text(size = 12, 
#             angle=90),
#         axis.ticks.x = element_line(linewidth = 0.5),
#         axis.ticks.y = element_line(linewidth = 0.5),
#         plot.title = element_text(size = 12),
#         legend.title = element_text(size = 12),
#         legend.text = element_text(size = 10),
#         legend.key.height = unit(1, 'cm'),
#         legend.key.width = unit(1, 'cm'),
#         legend.key.size = unit(1, 'cm'))
# 
# b1_m_scurve2 <- ggplot(error_scurve, 
#                      aes(x = bin1, 
#                          y = b_non_empty)) + 
#   geom_point(size = 1) +
#   geom_line(linewidth = 0.05) + 
#   # geom_vline(xintercept = 11, linetype="solid",
#   #            color = "#bdbdbd", linewidth=1) +
#   # geom_vline(xintercept = 9, linetype=2,
#   #            color = "#bdbdbd", linewidth=1) +
#   # geom_vline(xintercept = 13, linetype=2,
#   #            color = "#bdbdbd", linewidth=1) +
#   scale_x_continuous(breaks = 5:19) +
#   labs(x = expression(b[1]), y = expression(m)) +
#   ggtitle("(c)") +
#   theme(aspect.ratio = 0.75,
#         axis.text.x = element_text(size = 12),
#         axis.text.y = element_text(size = 12),
#         axis.title.x = element_text(size = 12),
#         axis.title.y = element_text(size = 12, 
#             angle=90),
#         axis.ticks.x = element_line(linewidth = 0.5),
#         axis.ticks.y = element_line(linewidth = 0.5),
#         plot.title = element_text(size = 12),
#         legend.title = element_text(size = 12),
#         legend.text = element_text(size = 10),
#         legend.key.height = unit(1, 'cm'),
#         legend.key.width = unit(1, 'cm'),
#         legend.key.size = unit(1, 'cm'))
```

@@

mse_scurve_b + mse_scurve_lwd + 
  mse_scurve_lwd2 + distance_hist + 
  plot_layout(ncol=2)
```
